# Functions and Iteration

## Automate creation of county-level median age graphs

### Write county-level median age graph function

Write a function to create a median age graph for a given state, based on two inputs: the 2 letter state abbreviation (used in the tidycensus call) and state name (used to label the graph)

```{r, warning=F, message=F}

# Write median age graph function
median_age_graph<-function(state_abbrev, state_full_name){
  census_call<-get_decennial(geography = "county",
                             state=state_abbrev,
                             variables = "P013001", 
                             geometry=TRUE,
                             year = 2010) %>% 
    rename(median_age=value) %>% 
    mutate(County=str_remove(NAME, paste0(" County,", " ", state_full_name))) %>% 
    select(-NAME)
  
  median_age_ggplot<-
    census_call %>%
    ggplot(aes(x = median_age, y = reorder(County, median_age))) + 
    geom_point()+
    labs(title=paste0("Median Age by County,", " ", state_full_name), x="Median Age", y="County")+
    theme(plot.title=element_text(hjust=0.5))
  
  return(median_age_ggplot)
}

# Test function for New Mexico
median_age_graph("NM", "New Mexico")

```

### Iteratively apply ```median_age_graph``` function to multiple input elements

Apply the ```median_age_graph``` function iteratively to New Mexico, Utah, Arizona, and Wyoming, thereby creating median age graphs for all these states; create a list to contain each of these output graphs as a separate list element

```{r}
# Create input vectors
state_codes<-c("NM", "UT", "AZ", "WY")
state_names<-c("New Mexico", "Utah", "Arizona", "Wyoming")

# Applies the "median_age_graph" function to multiple inputs (specified by the input vectors), and then places the outputs (i.e. median age graphs for New Mexico, Utah, Arizona, and Wyoming) in a list assigned to an object named "county_age_graphs_list"
county_age_graphs_list<-map2(.x=state_codes, .y=state_names, .f=median_age_graph)

# Open the list that contains the median age graphs
county_age_graphs_list

# Extract the Arizona graph by its index
county_age_graphs_list %>% pluck(3)

# Attach names to list elements and extract by name; for convenience, we'll name list elements based on "state_codes" vector
names(county_age_graphs_list)<-state_codes

# Extract the Wyoming graph by its name
county_age_graphs_list %>% pluck("WY")
```

## Automate creation of median-age maps

### Write function to create county-level median age maps

Write a function to create a county-level median age map for a given state, based on two inputs: the 2 letter state abbreviation (used in the tidycensus call) and state name (used to label the map)

```{r, fig.asp=0.5, message=F, warning=F}

# Write function to create county-level median age map for a given state 

median_age_map<-function(state_abbrev, state_full_name){
  census_call<-get_decennial(geography = "county",
                              state=state_abbrev,
                              variables = "P013001", 
                              geometry=TRUE,
                              year = 2010) %>% 
  rename(median_age=value) %>% 
  mutate(County=str_remove(NAME, paste0(" County,", " ", state_full_name))) %>% 
  select(-NAME)
  
  county_median_age_map<-
    tm_shape(census_call)+
    tm_polygons(col="median_age",
               style="jenks",
                palette="YlGnBu", 
                midpoint=TRUE)+
    tm_layout(frame=FALSE, 
              main.title=paste0("Median Age by County,\n", state_full_name),  
              main.title.position="left", 
              legend.outside=TRUE)
return(county_median_age_map)
}

# Test function for New Mexico
median_age_map("NM", "New Mexico")

```

### Iteratively apply ```median_age_map``` function to multiple input elements

```{r}
# Create input vectors
state_codes<-c("NM", "UT", "AZ", "WY")
state_names<-c("New Mexico", "Utah", "Arizona", "Wyoming")

# Applies the "median_age_map" function to multiple inputs (specified by the input vectors), and then places the outputs (i.e. median age maps for New Mexico, Utah, Arizona, and Wyoming) in a list assigned to an object named "county_age_maps_list"
county_age_maps_list<-map2(.x=state_codes, .y=state_names,median_age_map)

# Open the list that contains the median age maps
county_age_maps_list

# Extract the Utah map by its index
county_age_maps_list %>% pluck(2)

# Attach names (based on "state_codes" vector) to "county_age_maps_list" list elements 
names(county_age_maps_list)<-state_codes

# Extract the Arizona map by name
county_age_maps_list %>% pluck("AZ")
```







