# Advanced data retrieval

```{r}
population_rural_CO_UT<-get_decennial(geography = "tract",
                                      state=c("CO", "UT"),
                                      variables = c("P001001", "P002005"),
                                      output="wide",
                                      year = 2000) 
```


## Retrieving data from multiple census years

```{r, error=TRUE}
population_rural_2000_2010<-get_decennial(geography = "state", 
                                         variables = c("P001001", "P002005"),
                                         output="wide",
                                         year = c(2000,2010)) 
```


### Retrieving different census year datasets and placing them as separate elements in a list

```{r, message=F, warning=F}
my_years<-c(2000,2010)
census_query<-function(my_years){
  get_decennial(geography = "state", 
                  variables = c("P001001", "P002005"),
                  output="wide",
                  year=my_years) %>% 
    mutate(rural_pct=(P002005/P001001)*100) %>% 
    mutate(year=my_years)
}

population_rural_2000_2010_list<-map(my_years, census_query)

population_rural_2000_2010_list

population_rural_2000_2010_list %>% pluck(1)

names(population_rural_2000_2010_list)<-my_years

population_rural_2000_2010_list

population_rural_2000_2010_list %>% pluck("2010")

population_rural_2000_2010_list[["2010"]]

population_rural_2000_2010_list %>% pluck(2)

population_rural_2000_2010_list[[2]]
```


### Retrieving different census year datasets and placing them in a single data frame

```{r, message=F, warning=F}
population_rural_2000_2010_df<-map_dfr(my_years, census_query) %>% arrange(GEOID)

population_rural_2000_2010_df
```

## Retrieving microgeography data from multiple states

```{r, error=TRUE}
population_rural_CO_UT<-get_decennial(geography = "tract", 
                                      state=c("CO","UT"),
                                      variables = c("P001001", "P002005"),
                                      output="wide",
                                      year=2000) %>% 
                        mutate(rural_pct=(P002005/P001001)*100) %>% 
                        mutate(year=my_years)                 

```

### List

```{r}
my_states<-c("CO", "UT")
population_rural_CO_UT<-function(my_states){
  get_decennial(geography = "tract", 
                state=my_states,
                variables = c("P001001", "P002005"),
                output="wide",
                year=2010) %>% 
    mutate(rural_pct=(P002005/P001001)*100) %>% 
    mutate(Year=2010) %>% 
    mutate(STATE=my_states)
}


population_rural_CO_UT_list<-map(my_states, population_rural_CO_UT)
population_rural_CO_UT_list
population_rural_CO_UT_list %>% pluck(1)

names(population_rural_CO_UT_list)<-my_states

population_rural_CO_UT_list

population_rural_CO_UT_list %>% pluck("UT")
```

### df 

```{r}
population_rural_CO_UT_df<-map_dfr(my_states, population_rural_CO_UT)
population_rural_CO_UT_df
```

## Retrieving microgeography data from multiple states AND multiple years  

### List

```{r}
my_states<-c("CO", "CO", "UT", "UT")
my_years<-c(2000, 2010, 2000, 2010)

population_rural_CO_UT_2000_2010<-function(my_states, my_years){
  get_decennial(geography = "tract", 
                state=my_states,
                variables = c("P001001", "P002005"),
                output="wide",
                year=my_years) %>% 
    mutate(rural_pct=(P002005/P001001)*100) %>% 
    mutate(YEAR=my_years) %>% 
    mutate(STATE=my_states)
}

population_rural_CO_UT_2000_2010_list<-map2(my_states, my_years, population_rural_CO_UT_2000_2010)

population_rural_CO_UT_2000_2010_list

population_rural_CO_UT_2000_2010_list %>% pluck(2)

name_vector<-paste0(my_states, "_", my_years)

names(population_rural_CO_UT_2000_2010_list)<-name_vector

population_rural_CO_UT_2000_2010_list

population_rural_CO_UT_2000_2010_list %>% pluck("CO_2010")
```

### df

```{r}
population_rural_CO_UT_2000_2010_df<-map2_dfr(my_states, my_years, population_rural_CO_UT_2000_2010)
```

```{r}
population_rural_CO_UT_2000_2010_df
```




[https://mattherman.info/blog/tidycensus-mult-year/](https://mattherman.info/blog/tidycensus-mult-year/)





median age by county colorado
vis and map

functions and iteration

advanced retrieval

temporal dynamics

