# Advanced data retrieval

```{r}
# Use tidycensus to extract tract-level data from Colorado and Utah on population (P001001) and rural population (P002005) based on 2000 decennial census; assigns data to object named "population_rural_CO_UT"
population_rural_CO_UT<-get_decennial(geography = "tract",
                                      state=c("CO", "UT"),
                                      variables = c("P001001", "P002005"),
                                      output="wide",
                                      year = 2000) 
```


## Retrieving data from multiple census years

```{r, error=TRUE}
# Attempts to extract state level data (for all US states) on population (P001001) and rural population (P002005) based on 2000 decennial census and 2010 decennial census, and assign to new object named "population_rural_2000_2010"
population_rural_2000_2010<-get_decennial(geography = "state", 
                                         variables = c("P001001", "P002005"),
                                         output="wide",
                                         year = c(2000,2010)) 
```


### Retrieving different census year datasets and placing them as separate elements in a list

```{r, message=F, warning=F}

# Creates function that takes a given year as an input, and uses tidycensus to extract a state-level dataset of population (P001001) and rural population (P002005) for the specified year; the output dataset also includes new columns containing the state's rural population as a percentage of the overall population ("rural_pct") and the year to which the data corresponds ("year"); the function is assigned to a new object named "census_query"

census_query<-function(desired_year){
  get_decennial(geography = "state", 
                  variables = c("P001001", "P002005"),
                  output="wide",
                  year=desired_year) %>% 
    mutate(rural_pct=(P002005/P001001)*100) %>% 
    mutate(year=desired_year)
}

# Test the "census_query" object for the year 2000
census_query(2000)

```

```{r}
# Creates a new vector, assigned to an object named "my_years", containing the years for which we want to extract data
my_years<-c(2000,2010)

# Applies the "census_query" function to the inputs specified by the "my_years" vector, and deposits the two resulting datasets (one from 2000 and the other from 2010) as distinct elements in a list assigned to an object named population_rural_2000_2010_list
population_rural_2000_2010_list<-map(.x=my_years, .f=census_query)

# Prints the contents of "population_rural_2000_2010_list"
population_rural_2000_2010_list

# Extracts the first list element (the 2000 dataset) from the list using its index
population_rural_2000_2010_list %>% pluck(1)

# Names the list elements based on the input years (i.e. the years in "my_years")
names(population_rural_2000_2010_list)<-my_years

# prints "population_rural_2000_2010_list"; notice the list elements are now named after the years to which they correspond 
population_rural_2000_2010_list

# Extracts the 2010 dataset from "population_rural_2000_2010_list" using its newly assigned name
population_rural_2000_2010_list %>% pluck("2010")

# Note that even if a list element is named, you can still extract it using its index, rather than its name
population_rural_2000_2010_list %>% pluck(2)

```

### Retrieving different census year datasets and placing them in a single data frame

```{r, message=F, warning=F}
# Applies the "census_query" function to the inputs specified by the "my_years" vector, and deposits the resulting census datasets (one from 2000 and the other from 2010) into a single data frame assigned to an object named "population_rural_2000_2010_df"
population_rural_2000_2010_df<-map_dfr(.x=my_years, .f=census_query) %>% arrange(GEOID)

# prints the first few records of "population_rural_2000_2010_df" 
population_rural_2000_2010_df
```

## Retrieving microgeography data from multiple states

```{r, error=TRUE}
# Attempts to retrieve tract-level data on overall population (P001001) and rural population (P002005) from Colorado and Utah, based on the 2000 decennial census; then, creates a new variable named "rural_pct" that captures each tracts rural share of the overall population, as well as a new variable named "rural_pct"
population_rural_CO_UT<-get_decennial(geography = "tract", 
                                      state=c("CO","UT"),
                                      variables = c("P001001", "P002005"),
                                      output="wide",
                                      year=2000) %>% 
                        mutate(rural_pct=(P002005/P001001)*100)         

```

### Retrieving microgeography datasets for multiple states and placing datasets into a list

```{r}
# Creates a function assigned to an object named "population_rural_statetract" that takes a given state input ("desired_state), and passes this state into tidycensus's "get_decennial" function to retrieve a tract-level dataset for that state containing data on overall population (P001001) and rural population (P002005) based on the 2010 decennial census, while also creating three new variables: "rural_pct" (the percentage of the overall population that is rural), "Year" (a varaiable that takes on the value 2010, so that we can readily identify that the data corresponds to the year 2010), and "STATE" (an ID variable that will take on the value of the state input, so that we can readily identify the state for which the data has been extracted)
population_rural_statetract<-function(desired_state){
  get_decennial(geography = "tract", 
                state=desired_state,
                variables = c("P001001", "P002005"),
                output="wide",
                year=2010) %>% 
    mutate(rural_pct=(P002005/P001001)*100) %>% 
    mutate(Year=2010) %>% 
    mutate(STATE=desired_state)
}

# Test the function for a given state; here, California ("CA)
population_rural_statetract("CA")
```

Now that we know the function works, let's apply it to extract our desired dataset for Colorado and Utah, and place these datasets into a list:

```{r}
# define a vector of desired states and assign it to an object named "my_states"
my_states<-c("CO", "UT")

# Applies the "population_rural_state_tract" function to the inputs specified by the "my_states" vector, and deposits the two resulting datasets (one for CO, and the other for UT) as distinct elements in a list assigned to an object named population_rural_CO_UT_list
population_rural_CO_UT_list<-map(.x=my_states, .f=population_rural_statetract)

# Prints the contents of "population_rural_CO_UT_list"
population_rural_CO_UT_list

# Use the "pluck" function to extract the Colorado dataset from "population_rural_CO_UT_list" by its index
population_rural_CO_UT_list %>% pluck(1)

# Name the list elements in "population_rural_CO_UT_list" based on the "my_states" vector
names(population_rural_CO_UT_list)<-my_states

# Print "population_rural_CO_UT_list" list contents; notice that they're now named based on the state
population_rural_CO_UT_list

# Use the "pluck" function to extract the Utah dataset from "population_rural_CO_UT_list" by name
population_rural_CO_UT_list %>% pluck("UT")
```

### Retrieving microgeography data from multiple states and placing them into the same data frame

```{r}
# Applies the "population_rural_statetract" function to the inputs specified by the "my_states" vector, and deposits the resulting census datasets (one CO and the other UT) into a single data frame assigned to an object named "population_rural_CO_UT_df"
population_rural_CO_UT_df<-map_dfr(.x=my_states, .f=population_rural_statetract)

# Prints the contents of "population_rural_CO_UT_df"
population_rural_CO_UT_df
```

## Retrieving microgeography data from multiple states AND multiple years

We can use the same principles to extract census datasets at sub-state geographies for multiple states, across from multiple years. 

### Retrieving data from multiple years, for multiple states, and placing the datasets from the various state/year combinations into a list

```{r}
# Define a function (assigned to an object named "population_rural_statetract_year") that takes a state input ("desired_state") and a year input ("desired_year), and returns a tract level dataset for that state and census year which contains variables specifying the total population, and the rural population, while also creating three new variables: "rural_pct" (the percentage of the overall tract population that is rural), "YEAR" (a varaiable that takes on the value of the input, so that we can readily identify what year the data corresponds to), and "STATE" (an ID variable that will take on the value of the state input, so that we can readily identify the state for which the data has been extracted)

population_rural_statetract_year<-function(desired_state, desired_year){
  get_decennial(geography = "tract", 
                state=desired_state,
                variables = c("P001001", "P002005"),
                output="wide",
                year=desired_year) %>% 
    mutate(rural_pct=(P002005/P001001)*100) %>% 
    mutate(YEAR=desired_year) %>% 
    mutate(STATE=desired_state)
}

# Tests the "population_rural_CO_UT_2000_2010" function for Virginia, for the 2010 census
population_rural_statetract_year("VA", 2010) 

```

Now, we can iteratively apply the ```population_rural_statetract_year``` function to multiple state/year inputs. In the example below, we use ```population_rural_statetract_year``` to extract data from the 2000 and 2010 censuses, for both Colorado and Utah, and deposit the resulting four datasets (CO/2000, CO/2010, UT/2000, UT/2010) into a list

```{r}
# define a vector of desired states to iterate over, and assign it to an obejct named "my_states"
my_states<-c("CO", "CO", "UT", "UT")

# define a vector of desired years to iterate over, and assign it to an object named "my_years"
my_years<-c(2000, 2010, 2000, 2010)

# Applies the "population_rural_statetract_year" function to the inputs specified by the "my_states" and "my_year" vectors, and deposits the resulting four census datasets (one for CO 2000; one for CO 2010; one for UT 2000; one for UT 2010)  as distinct elements in a list that is assigned to an object named "population_rural_CO_UT_2000_2010_list"
population_rural_CO_UT_2000_2010_list<-map2(.x=my_states, .y=my_years, .f=population_rural_statetract_year)

# Prints contents of "population_rural_CO_UT_2000_2010_list"
population_rural_CO_UT_2000_2010_list

# Uses the ```pluck``` function to extract the Colorado/2010 dataset from "population_rural_CO_UT_2000_2010_list" 
population_rural_CO_UT_2000_2010_list %>% pluck(2)

# Creates a name vector in which the four different state/year combinations are connected by an underscore ("_") and assigns the resulting vector to an object named "name_vector"
name_vector<-paste0(my_states, "_", my_years)

# Prints "name_vector" to inspect contents
name_vector

# Assigns the names in "name_vector" to the list elements
names(population_rural_CO_UT_2000_2010_list)<-name_vector

# Prints the contents of "population_rural_CO_UT_2000_2010_list"; note that the list elements are now named according to the state/year inputs upon which they are based
population_rural_CO_UT_2000_2010_list

# Uses the ```pluck``` function to extract the 2010 Colorado dataset by name
population_rural_CO_UT_2000_2010_list %>% pluck("CO_2010")
```

### Retrieving data from multiple years, for multiple states, and placing all of the data into a single data frame

```{r}
# # Applies the "population_rural_statetract_year" function to the inputs specified by the "my_states" and "my_years" vectors, and deposits the resulting four census datasets (CO/2000, CO/2010, UT/2000, UT/2010) into a single data frame assigned to an object named "population_rural_CO_UT_2000_2010_df"
population_rural_CO_UT_2000_2010_df<-map2_dfr(.x=my_states, .y=my_years, .f=population_rural_statetract_year)
```

```{r}
# Prints the contents of "population_rural_CO_UT_2000_2010_df" 
population_rural_CO_UT_2000_2010_df
```




[https://mattherman.info/blog/tidycensus-mult-year/](https://mattherman.info/blog/tidycensus-mult-year/)





median age by county colorado
vis and map

functions and iteration

advanced retrieval

temporal dynamics

