# (PART) American Community Survey (ACS) {-}

# ACS Data Retrieval

## Extracting variable codes for ACS data

To inspect the variable list for the ACS, use the “load variables” function. Let’s say we want to work with the 5-year ACS ending in 2019:

```{r}
ACS_5_2019<-load_variables(2019,"acs5")
```

```{r}
ACS_5_2019
```

If we wanted to view the variables for the 5-year acs ending in 2015, we'd use 

```{r}
ACS_5_2015<-load_variables(2015,"acs5")
ACS_5_2015
```

If we wanted to view the variables for the 1-year acs ending in 2017, we'd use:

```{r}
ACS_1_2017<-load_variables(2017, "acs1")
ACS_1_2017
```

3-year ACS ending in 2012:

```{r}
ACS_3_2012<-load_variables(2012, "acs3")
ACS_3_2012
```

## Calling ACS data with ```tidycensus``` 

County-level median income data

```{r, warning=F, message=F}
median_income<-get_acs(geography="county",
                       variables="B19013_001",
                       year=2019) %>% 
              rename(median_income=estimate) %>% 
              arrange(desc(median_income))
```

```{r}
median_income
```







