# Exploratory analysis and visualization with ACS data 

## Identifying and visualizing the counties from each state with the highest median income

```{r}
highest_income_counties<-median_income %>% 
                          separate(NAME,c("County","State"),sep=",") %>% 
                          group_by(State) %>% 
                          arrange(desc(median_income)) %>% 
                          slice_max(median_income, n=1) %>% 
                          unite(NAME, c("County","State"), remove=FALSE, sep=",")

highest_income_counties

```
 
```{r}
highest_income_counties_viz<-
  highest_income_counties %>% 
        ggplot(aes(x=median_income,y=reorder(NAME, median_income)))+
        geom_errorbarh(aes(xmin = median_income - moe, xmax = median_income + moe)) +
        geom_point(color = "red", 
                   size = 3)+
        labs(title="County with Highest Median Income, by State",
             y="",
             x="Median Income Estimate from 5-year ACS\n(bars indicate margin of error)")+ 
        theme(plot.title=element_text(hjust=0.5))

```

```{r, fig.width=9,fig.height=8}
highest_income_counties_viz
```

## Visualizing county-level median income for Colorado

```{r, fig.width=9,fig.height=8}
median_income_CO_counties_2018<-get_acs(geography="county",
                                     state="CO",
                                     variables="B19013_001",
                                     year=2018) %>% 
                            rename(median_income=estimate) %>% 
                            arrange(desc(median_income))

median_income_CO_counties_2018_viz<-
     median_income_CO_2018 %>% 
        mutate(County_Name=str_remove_all(NAME,"County, Colorado")) %>% 
          ggplot(aes(x=median_income,y=reorder(County_Name, median_income)))+
           geom_errorbarh(aes(xmin = median_income - moe, xmax = median_income + moe)) +
           geom_point(color = "blue", size = 3)+
                      labs(title="Median Income in Colorado, by County (2018)",
                              y="", x="Median Income Estimate from 5 year ACS\n(Bars indicate margin of error)")+
                             theme(plot.title=element_text(hjust=0.5))

```

```{r}
median_income_CO_counties_2018_viz  
```

## Writing functions to visualize ACS data

```{r}

county_median_income<-function(state_abbrev, state_name){
  median_income<-get_acs(geography="county",
                         state=state_abbrev,
                         variables="B19013_001",
                         year=2018) %>% 
    rename(median_income=estimate) %>% 
    arrange(desc(median_income))
  
  string_to_remove<-paste0("County", ", ", state_name)
  chart_title<-paste0("Median Income in ", state_name, ", by County (2018)")
  
  median_income_viz<-median_income %>% 
                      mutate(County_Name=str_remove_all(NAME,string_to_remove)) %>% 
                      ggplot(aes(x=median_income,y=reorder(County_Name, median_income)))+
                      geom_errorbarh(aes(xmin = median_income - moe, xmax = median_income + moe)) +
                      geom_point(color = "blue", size = 3)+
                      labs(title=chart_title,
                           y="", x="Median Income Estimate from 5 year ACS\n(Bars indicate margin of error)")+
                      theme(plot.title=element_text(hjust=0.5))+
                      scale_x_continuous(labels=scales::comma)
  return(median_income_viz)
}
```

```{r, fig.width=9,fig.height=8}
# test function 
county_median_income("NY", "New York")
```

```{r}
# Creates vector of state abbreviations
state_abbrev_vector<-c("NM", "UT", "AZ", "WY")
# Creates vector of state names
state_name_vector<-c("New Mexico", "Utah", "Arizona", "Wyoming")

# Iteratively applies "county_median_income" function to inputs specified by "state_abbrev" vector
# and "state_name" vector, and assigns resulting visualization outputs to list named "median_income_2018_viz_list"
median_income2018_viz_list<-map2(state_abbrev_vector, state_name_vector, county_median_income)

```

```{r}
# Creates a vector of names for "median_income2018_viz_list" list elements
names_median_income_list<-paste0(state_abbrev_vector, "_", "2018_ACS_medianincome")
# Inspect names in "names_median_income_list" vector
names_median_income_list
```

```{r}
# Assigns names in 
names(median_income2018_viz_list)<-names_median_income_list
```

```{r, fig.width=9,fig.height=8, warning=F}
median_income2018_viz_list
```

## Dynamic mapping

```{r}
median_income_2016_states<-get_acs(geography="state",
                                     variables="B19013_001",
                                     year=2016,
                                     geometry=TRUE) %>% 
                            rename(median_income=estimate) %>% 
                            arrange(desc(median_income)) %>% 
                            relocate(NAME)
```

```{r}
median_income_map<-tm_shape(median_income_2016_states)+
                   tm_polygons(col="median_income", n=6, style="fisher", palette="YlOrBr")
tmap_mode("view")

median_income_map
```

# Bivariate Analysis with ACS Data: Education and Income

What is the relationship between county median income, and county share of the over-25 population without at least a bachelor's degree? 

First, let's inspect the variable codes for the 2019 ACS:

```{r}
ACS_5_2018<-load_variables(2018, "acs5")
```

```{r}
education_variables<-ACS_5_2018 %>% filter(str_detect(name, 'B15003'))
```

```{r}
View(education_variables)
```

```{r}
kable(education_variables)
```

```{r}
education_vars<-c(Bachelors="B15003_022", Masters="B15003_023", Professional="B15003_024", 
                  Doctorate="B15003_025") 

education_acs_2018<-get_acs(geography="county",
                            year=2018,
                            variables=education_vars,
                            summary_var="B15003_001")
```

```{r}
education_acs_2018
```

```{r}
pct_less_than_BA<-education_acs_2018 %>% 
                  group_by(GEOID, NAME) %>% 
                  mutate(no_degree_pct=((summary_est-sum(estimate))/(summary_est)*100)) %>% 
                  summarize(mean(no_degree_pct)) %>% 
                  rename(no_degree_pct="mean(no_degree_pct)")

pct_less_than_BA
```



```{r}
median_income_2018<-get_acs(geography="county",
                       variables="B19013_001",
                       year=2018) %>% 
                rename(median_income=estimate) %>% 
                arrange(desc(median_income))
```

```{r}
median_income_nodegree<-full_join(median_income_2018, pct_less_than_BA,by="GEOID")
```


```{r}
median_income_nodegree
```

```{r}
medianincome_nodegree_viz<-median_income_nodegree %>% 
                           ggplot()+
                           geom_point(aes(x=no_degree_pct,y=median_income))+
                           geom_smooth(aes(x=no_degree_pct,y=median_income),method="lm")+
                           ylab("Estimated County Median Income")+
                           xlab("Estimated Percentage of Under-25 Population Without at least Bachelors Degree")
```


```{r}
medianincome_nodegree_viz
```

```{r}
median_income_nodegree<-median_income_nodegree %>% 
                        separate(NAME.x,c("County","State"),sep=",")

median_income_nodegree
```

```{r}
medianincome_nodegree_statefacets_viz<-median_income_nodegree %>% 
                                   ggplot()+
                                   geom_point(aes(x=no_degree_pct,y=median_income))+
                                   geom_smooth(aes(x=no_degree_pct,y=median_income),method="lm")+
                                   ylab("County Median Income")+
                                   xlab("Pct No Degree")+
                                   facet_wrap(~State)

```

```{r}
medianincome_nodegree_statefacets_viz
```


Nested dataset

```{r}
median_income_nodegree_nested<-median_income_nodegree %>% 
  group_by(State) %>% 
  nest() %>% 
  arrange(State)

median_income_nodegree_nested
```

```{r}
median_income_nodegree_nested %>% pull(2) %>% pluck(1)
```

```{r}
median_income_nodegree_nested %>% pull("data") %>% pluck(6)

```

Add median income range

```{r}
median_income_range<-function(x){
  max_median_income<-x %>% slice_max(median_income, n=1)
  min_median_income<-x %>% slice_min(median_income, n=1)
  range<-max_median_income$median_income-min_median_income$median_income
  return(range)
}

median_income_nodegree_nested<-median_income_nodegree_nested %>% 
  mutate(county_income_range=map_dbl(data, median_income_range))

```

```{r}
median_income_nodegree_nested
```


```{r}
# Five states with highest spatial inequality with respect to county median income
median_income_nodegree_nested %>% ungroup() %>% slice_max(county_income_range, n=5)
```


```{r}
# Five states with lowest spatial inequality with respect to county median income
median_income_nodegree_nested %>% ungroup() %>% slice_min(county_income_range, n=5)
```


```{r}
correlation_income_ed<-function(x){
  income<-x$median_income
  education<-x$no_degree_pct
  corr_table<-cor.test(income, education)
  return(corr_table)
}

median_income_nodegree_nested_final<- 
  median_income_nodegree_nested %>% 
  filter(county_income_range>0) %>% 
  mutate(correlation_education_income=map(data, correlation_income_ed))

```

```{r}
median_income_nodegree_nested_final
```

```{r}
# Alabama correlation coefficient table
median_income_nodegree_nested_final %>% pull(4) %>% pluck(1)
```

```{r}
# Colorado correlation coefficient table
median_income_nodegree_nested_final %>% pull(4) %>% pluck(6)
```

# Integrating census data and external data
